\LARGE \bf Exploiting Non-Slip Wall Contacts \\ to Position Two Particles Using The Same Control Input

### 2018-05-22 journalWallFriction/experiment.tex
A lot of changes to figures.
Only requires the algorithm for irregular polygon

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Results}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To obtain experimental data, the test samples (the phantom intestine model and the bovine cross section) were placed in laser-cut acrylic discs and then immersed in corn syrup. Corn syrup was used to increase the viscosity to 12000 cP for the experiments. Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles. The magnetic field in this setup is only approximately uniform. The magnetic force varies in both magnitude and orientation. As shown in the video attachment, the particle closer to the coil moves faster than the other particle. Also, magnetic forces are not exactly parallel, but point toward the activated coil. Algorithm~\ref{alg:optimalAlg} is robust to these non uniformities, but sometimes requires additional iterations.

@@

Strips of cow stomach approximately $\SI{5}{\milli\metre}$ thick were cut and sewn to acrylic cylinder and then glued to an acrylic substrate using cyanoacrylate (superglue). This assemble was then filled with corn syrup. The experiment is shown in Fig.~\ref{fig:story} bottom row. 
%In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines. This meant more care had to be taken in navigating the magnets to their goal locations to avoid getting them too close to each other.

@@

\begin{figure*}\label{fig:3dPrinted}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
\\
\vspace{1em}
\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
\caption{\label{fig:story}
Frames showing particle positions before and after control inputs. Top row: small intestine phantom. Bottom row: cow's stomach tissue.
} \vspace{-1em}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Results}\label{sec:expResults}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[htb!]\label{fig:3dPrinted}
\centering
\vspace{1.5em}
%\begin{overpic}[width=\columnwidth]{firstImage.jpg}\end{overpic}
\begin{overpic}[width=2\columnwidth]{3dexperiment.pdf}\end{overpic}
\\
\vspace{1em}
\begin{overpic}[width=2\columnwidth]{realTripe.pdf}\end{overpic}
\caption{\label{fig:story}
Frames showing particle positions before and after control inputs. Top row: small intestine phantom. Bottom row: cow stomach tissue.
} \vspace{-1em}
\end{figure*}

@@

To obtain experimental data, the test samples (the phantom intestine model and the bovine cross section) were placed in laser-cut acrylic discs and then immersed in corn syrup. Corn syrup was used to increase the viscosity to 12000 cP for the experiments. Spherical $\SI{1}{\milli\metre}$ magnets (supermagnetman \#SP0100-50) were used as our particles. The magnetic field in this setup is only approximately uniform. The magnetic force varies in both magnitude and orientation. As shown in the video attachment, the particle closer to the coil moves faster than the other particle. Also, magnetic forces are not exactly parallel, but point toward the center of activated coil. Algorithm~\ref{alg:optimalAlg} is robust to these non-uniformities, but sometimes requires additional iterations.

@@

Strips of cow stomach approximately $\SI{5}{\milli\metre}$ thick were cut and sewn to acrylic cylinder and then glued to an acrylic substrate using cyanoacrylate (superglue). This assembly was then filled with corn syrup. The experiment is shown in Fig.~\ref{fig:story} bottom row. 
%In using the small intestines, which were about 25mm in diameter, the resulting workspace was less than half of that of our simulated intestines. This meant more care had to be taken in navigating the magnets to their goal locations to avoid getting them too close to each other.

@@

### 2018-05-22 journalWallFriction/intro.tex
Only requires the algorithm for irregular polygon

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{subsec:WallFriction} introduces a  model for boundary interaction.  
We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

This paper is elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

>>>>

The paper is arranged as follows. 
After a review of recent related work in Sec.  \ref{sec:RelatedWork},
 %Sec. \ref{subsec:FluidInTank} provides analytical position control results of stable configurations in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes,
  Sec.  \ref{sec:theory} introduces a  model for boundary interaction and two shortest path results for representative workspaces.   
We provide a shortest-path algorithm to arbitrarily position two particles in Sec.  \ref{sec:PostionControl2Robots}.
Section  \ref{sec:simulation} describes implementations of the algorithms in simulation and  Sec.  \ref{sec:expResults} describes hardware experiments, as shown in Fig.~\ref{fig:IntroPic}. 
 We end with directions for future research in Sec.  \ref{sec:conclusion}.

This paper is an elaboration of preliminary work in a conference paper \cite{shahrokhi2017algorithms} which considered only square workspaces. This work extends the analysis to convex workspaces and 3D positioning. This paper also implements the algorithms using a hardware setup inspired by the anatomy of the gastrointestinal tract.

### 2018-05-22 journalWallFriction/simulation.tex
Only requires the algorithm for irregular polygon

%\subsection{Position Control of Two Robots}

Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. 
%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
 %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.

 The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings. Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} and middle and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.5,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace. Fig.~\ref{fig:contour} right shows the total distance of the path. This plot shows the nonlinear nature of the path planning. The path length grows when the goals have $\pi$ difference and are very close to the boundary. 
 The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance). This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}. 
 Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace. Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles. All of the discussed plots have considered the particles to be unique. If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller. The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.

@@

\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\end{overpic}
\caption{\label{fig:contour}
Plots show performance with one goal on the boundary.
}
\end{figure}

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
%\vspace{1em}
\begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
\caption{\label{fig:deltanumdist}
The worst case occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
} 
\vspace{-1em}
\end{figure}

\end{overpic}
\begin{overpic}[width =\figwid]{identical.png}\put(-2,6){\begin{turn}{90} \tiny{interchangeable particles}
\end{turn}}
\end{overpic}
}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
 The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
\vspace{-1em}
}
\end{figure}

>>>>

%\subsection{Position Control of Two Robots}
\begin{figure}
\centering
\begin{overpic}[width=0.49\columnwidth]{worstnum.pdf}\put(0,75){a)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{worstdist.pdf}\put(0,75){b)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoalnum.pdf}\put(0,75){c)}\end{overpic}
\begin{overpic}[width=0.49\columnwidth]{middlegoaldist.pdf}\put(0,75){d)}\end{overpic}
\caption{\label{fig:contour}
Plots show performance with one goal on the boundary.
}
\end{figure}

\begin{figure}
\centering
%\begin{overpic}[width=\columnwidth]{deltanum.pdf}\end{overpic}\\
%\vspace{1em}
\begin{overpic}[width=\columnwidth]{deltadist.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:deltanumdist}
The worst case path length occurs when particles must swap antipodes. This can never be achieved but can be asymptotically approached. Plot shows decreasing error as the number of moves grows.
} 
\end{figure}

\end{overpic}
\begin{overpic}[width =\figwid]{identical.png}\put(-2,6){\begin{turn}{90} \tiny{interchangeable particles}
\end{turn}}
\end{overpic}
}\caption{\label{fig:contourPlots}{Starting positions of particles $1$ and $2$ and goal position of particle $2$ are fixed, and $\epsilon=0.001$.
 The top row of contour plots show the distance if robot $1$'s goal position is varied in $x$ and $y$. The bottom row shows the number of moves required for the same configurations.}
\vspace{-1em}
}
\end{figure}
Algorithm \ref{alg:optimalAlg}  was implemented in Mathematica using particles with zero radius. 
%An online interactive demonstration and source code of the algorithm are available at \cite{Shahrokhi2015mathematicaParticle}.
%  Fig.~\ref{fig:shapeControlMathematica1}  shows  an implementation of this algorithm with robot initial positions represented by hollow squares and final positions by circles. 
 %Dashed lines show the shortest route if robots could be controlled independently, while solid lines show the optimal shortest  path using uniform inputs.

 The contour plots in Fig.~\ref{fig:contour} left show the length of the path for two different settings. Top row considers \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(-0.2,0)$\} and bottom row considers  \{$s_1,s_2,g_1$\} = \{$(0.2,0.2),(-0.1,-0.1),(0,0)$\} in a workspace with $r= 0.5$, and $g_2$ ranging over all the workspace. Fig.~\ref{fig:contour} right shows the total distance of the path. This plot shows the nonlinear nature of the path planning. The path length grows when the goals have $\pi$ difference and are very close to the boundary. TODO:???? We need to discuss the new plots.
 The worst case occurs when the ending points are at antipodes along the boundary ($\pi$ angular distance). This can never be achieved but can be asymptotically approached as shown in Fig.~\ref{fig:deltanumdist}. 
 Fig.~\ref{fig:contourPlots} shows the same concepts in a square workspace. Fig.~\ref{fig:contourPlots} top and middle row considers the particles for three arbitrary starting and goal positions for the particles. All of the discussed plots have considered the particles to be unique. If particles are interchangeable and it does not matter to put the specific particle to its goal location, the path length will be significantly smaller. The bottom row of  Fig.~\ref{fig:contourPlots} considers interchangeable particles with the same configuration as the middle row with unique particles.

@@

### 2018-05-22 journalWallFriction/theory.tex
Only requires the algorithm for irregular polygon

%  This equation can be simplified to 
% 
% \begin{equation}
%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
% \end{equation}
% 
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points (blue square) to (blue ellipse) in the unit disk that intersects the circumference. The path length as a function of intersection point, $(\cos\theta,\sin\theta)$ is shown at right. See \cite{BeckerShortestPath}.}
%\vspace{-1em}
}
\end{figure}

 The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}. There are several simple solutions. If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0. If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$. Label the origin $O$. 
 The optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
 The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
 We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is exactly $P$. 
  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
 From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$. If we mirror the point $S$ about line $\overline{OP}$ and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.

 Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$. We evaluate each and select the solution that results in the shortest length path. This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection. For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}. The closed form solution is long, and is in the paper attachments.

>>>>

%  This equation can be simplified to 
% 
% \begin{equation}
%\sqrt{1+s^2-2 s \cos \theta}+  \sqrt{1+e^2-2 e \cos(\beta-\theta)}. 
% \end{equation}
% 
\begin{figure}
\centering
\renewcommand{\figwid}{\columnwidth}
{\begin{overpic}[width =\figwid]{shortestpath.pdf}
\end{overpic}
}
\caption{\label{fig:shortestpath}{The shortest path between two points $S$ to $E$ in the unit disk that intersects the circumference. The path length as a function of intersection point, $P= (\cos\theta,\sin\theta)$ is shown at right. See \cite{BeckerShortestPath}.}
%\vspace{-1em}
}
\end{figure}

 The length of the two line segments as a function of $\theta$ is drawn in the right plot of Fig.~\ref{fig:shortestpath}. There are several simple solutions. If $s$ is 1 or $e$ is 0 or $\beta$ is 0, the optimal angle $\theta^*$ is 0. If $e$ is 1 or $s$ is 0, the optimal angle is $\beta$. Label the origin $O$. 
 The optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection.
 The angle $\angle{OPS}$ (from the origin to $P$ to $S$) is the same as the angle $\angle{OPE}$ (from the origin to $P$ to $E$). 
 We name these angles $\alpha$. This can be proved by drawing an ellipse whose foci are $S$ and $E$. When the ellipse is tangent to the circle, the point of tangency is $P$. 
  Since the distance from the origin to $P$ is always 1, we can set up three equalities using the law of sines:
 From triangle $OSP$: $\frac{\sin \alpha}{s}=\frac{\sin(\alpha + \theta)}{1}=\frac{\sin \theta}{||SP||}$, and from triangle $OEP$: $\frac{\sin \alpha}{e}=\frac{\sin(\beta - \theta)}{||EP||}$. If we mirror the point $S$ about line $\overline{OP}$ and label this point $C$, from triangle $CEO$: $\frac{\sin(\alpha + \theta)}{e}=\frac{\sin(2 \theta - \beta)}{||CE||}$.

 Simplifying this system of equations results in $s=e \csc \theta (s \sin(2 \theta-\beta)+\sin(\beta-\theta))$. Solving this last equation results in a quartic solution that has a closed-form solution with four roots, each of which can be either a clockwise or a counterclockwise rotation $\theta$, depending on the sign of $\beta$, with $-\pi\leq\beta\leq\pi$. We evaluate each and select the solution that results in the shortest length path. %This optimal path satisfies the law of reflection off the unit circle, with angle of incidence equal to angle of reflection. 
 For an interactive Mathematica demonstration of this shortest path, see \cite{BeckerShortestPath}. Because the closed form solution is long, it is included in the paper attachments.
